State-Space Model to Estimate Salmon Escapement Using Multiple Data Sources

### 2020-09-05 analysis/paper/paper.Rmd
incorporated new STADEM results and changed results on CV of escapement estimates in manuscript.

---
title: "State-Space Model to Estimate Salmon Escapement Using Multiple Data Sources"
author:
  - Kevin E. See:
      email: Kevin.See@merck.com
      institute: [BM]
      correspondence: true
  - Ryan N. Kinzer:
      email: ryank@nezperce.org
      institute: [NezPerce]
      correspondence: false
  - Michael W. Ackerman:
      email: mike.ackerman@merck.com
      institute: [BM]
      correspondence: false
institute:
  - BM: Biomark, Inc. 705 South 8th St., Boise, Idaho, 83702, USA
  - NezPerce: Nez Perce Tribe, Department of Fisheries Resource Management, 14054 Burr Road, PO Box 1942, McCall, Idaho, 83638, USA
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
    bookdown::pdf_document2:
      fig_caption: yes
      number_sections: FALSE
      toc: no
      geometry: margin=1in
      latex_engine: "xelatex"
      fig_width: 7
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
      includes:
        in_header: ../templates/header_manuscript.tex
    bookdown::html_document2:
      fig_caption: yes
      number_sections: FALSE
      toc: no
      fig_width: 7
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      number_sections: FALSE
      toc: no
      always_allow_html: true
      fig_width: 7
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
fontsize: 12pt
mainfont: Times New Roman
bibliography:
- references.bib
- packages.bib
- AckermanLibrary.bib
csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
abstract: |
  Accurate estimates of the number of salmonids passing Lower Granite Dam on the Snake River, by species and origin, are a critical input to assessing the status and trends of various populations, as well as successfully management of fisheries in that area. Here we describe a state-space model that estimates such escapement past a dam by using window counts, PIT tag observations and data from an adult fish trap, accounting for issues such as night-time passage, fallback and re-ascension, potential observation error at the window and uncertainty in the adult trap rate. We tested this approach using a simulation framework that mimicked several levels of observation error, differences between night-time passage and re-ascension rates and the possibility of the adult trap being closed for some period of time, and found it to produce unbiased estimates across all tested scenarios. We also applied this model to data from the Lower Granite dam to produce estimates of wild, clipped hatchery and unclipped hatchery spring-summer run Chinook Salmon and steelhead from spawn years 2010-2019. 
keywords: |
  escapement; Chinook; steelhead; salmon; Lower Granite Dam; state-space model; Bayesian model
highlights: |
  These are the highlights. 
---

>>>>

---
title: "State-Space Model to Estimate Salmon Escapement Using Multiple Data Sources"
author:
  - Kevin E. See:
      email: Kevin.See@merck.com
      institute: [BM]
      correspondence: true
  - Ryan N. Kinzer:
      email: ryank@nezperce.org
      institute: [NezPerce]
      correspondence: false
  - Michael W. Ackerman:
      email: mike.ackerman@merck.com
      institute: [BM]
      correspondence: false
institute:
  - BM: Biomark, Inc. 705 South 8th St., Boise, Idaho, 83702, USA
  - NezPerce: Nez Perce Tribe, Department of Fisheries Resource Management, 14054 Burr Road, PO Box 1942, McCall, Idaho, 83638, USA
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
    bookdown::pdf_document2:
      fig_caption: yes
      number_sections: FALSE
      toc: no
      geometry: margin=1in
      latex_engine: "xelatex"
      fig_width: 7
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
      includes:
        in_header: ../templates/header_manuscript.tex
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
      number_sections: FALSE
      toc: no
      always_allow_html: true
      fig_width: 7
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
    bookdown::html_document2:
      fig_caption: yes
      number_sections: FALSE
      toc: no
      fig_width: 7
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks2.lua
      - --lua-filter=../templates/pagebreak.lua
fontsize: 12pt
mainfont: Times New Roman
bibliography:
- references.bib
- packages.bib
- AckermanLibrary.bib
csl: "../templates/american-fisheries-society.csl" # Insert path for the bib-style
abstract: |
  Accurate estimates of the number of salmonids passing Lower Granite Dam on the Snake River, by species and origin, are a critical input to assessing the status and trends of various populations, as well as successfully management of fisheries in that area. Here we describe a state-space model that estimates such escapement past a dam by using window counts, PIT tag observations and data from an adult fish trap, accounting for issues such as night-time passage, fallback and re-ascension, potential observation error at the window and uncertainty in the adult trap rate. We tested this approach using a simulation framework that mimicked several levels of observation error, differences between night-time passage and re-ascension rates and the possibility of the adult trap being closed for some period of time, and found it to produce unbiased estimates across all tested scenarios. We also applied this model to data from the Lower Granite dam to produce estimates of wild, clipped hatchery and unclipped hatchery spring-summer run Chinook Salmon and steelhead from spawn years 2010-2019. 
keywords: |
  escapement; Chinook; steelhead; salmon; Lower Granite Dam; state-space model; Bayesian model
highlights: |
  These are the highlights. 
---
